---
output: 
  md_document:
    variant: gfm
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "./man/figures/README-",
  out.width = "100%",
  message = FALSE,
  warning = FALSE
)
```

# doestyle

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![R-CMD-check](https://github.com/nsw-education/doestyle/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/nsw-education/doestyle/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

`{doestyle}` is an R package that helps produce brand-compliant figures and tables for NSW Department of Education publications. It is under development, but already includes features that ease the use of brand colours and typesetting in `ggplot2` figures and `flextable` tables.

If you're interested in contributing to development, please [read the contributors' guide](./CONTRIBUTING.md) and/or watch [contributing to `doestyle`](https://schoolsnsw.sharepoint.com/:v:/s/DataVisWorkingGroup/EYthJPlfW4ZOhP-6aT0JrMABgWWB2bX_sEOVSlPMwlZwWg?e=OxCP6i&nav=eyJyZWZlcnJhbEluZm8iOnsicmVmZXJyYWxBcHAiOiJTdHJlYW1XZWJBcHAiLCJyZWZlcnJhbFZpZXciOiJTaGFyZURpYWxvZy1MaW5rIiwicmVmZXJyYWxBcHBQbGF0Zm9ybSI6IldlYiIsInJlZmVycmFsTW9kZSI6InZpZXcifX0%3D).

## Installation

Install the latest official release of `doestyle` with `install.packages()`:

```r
install.packages('doestyle', repos = c('https://nsw-education.r-universe.dev', 'https://cloud.r-project.org'))
```

Alternatively, install the latest in-development version from GitHub with `remotes`:

``` r
# Install `remotes` if needed
if (!require("remotes")) install.packages("remotes")

# Install doestyle using `remotes`
remotes::install_github("nsw-education/doestyle")
```

## Usage examples

The following examples use data from the [Public Schools Master Dataset](https://data.nsw.gov.au/data/dataset/nsw-education-nsw-public-schools-master-dataset), which is available as a package dataset named `public_schools`.

```{r}
library(tidyverse)
library(doestyle)

# The Public Schools Master Dataset
head(public_schools)
```

### Quick branding

`doestyle` provides the scale functions `scale_colour_doe()` and `scale_fill_doe()` to easily apply brand colours to the colour or fill aesthetics of a ggplot2 object.

For a default colour scale, apply `scale_colour_doe()` with no arguments:

```{r fig-lbote-foei-iec-scatter}
public_schools |>
  filter(Operational_directorate == "Metropolitan South and West") |>
  ggplot(aes(x = LBOTE_pct, y = FOEI_Value, colour = Intensive_english_centre)) +
  geom_point() +
  # Add a default department scale to the `colour` aesthetic. Note that `colour`
  # has already been mapped onto the value of `Intensive_english_centre` in this
  # plot's aesthetic mapping.
  scale_colour_doe()
```

An on-brand default fill scale can be applied with `scale_fill_doe()`. Be aware that lighter colours, although part of the department's official brand palette, may require a darker edge to meet [our accessibility requirements for contrast](https://brand.education.nsw.gov.au/content/edam/en/brand-guideline.html). 


```{r fig-connected-communities}
public_schools |>
  filter(str_detect(Principal_network, "Connected Communities")) |>
  ggplot(aes(x = Principal_network, fill = Level_of_schooling)) +
  geom_bar(colour = "black", position = position_dodge(preserve = "single")) +
  theme(legend.position = "bottom") +
  labs(y = "Schools") +
  # Add a default department scale to the `fill` aesthetic. Note that `fill` has
  # already been mapped onto the value of `Level_of_schooling` in this plot's
  # aesthetic mapping.
  scale_fill_doe()
```

### Colours

`doestyle` provides a mapping from the NSW Government named colour palette to hex values that can be used in almost any R function that handles colours. Every colour defined by the [NSW Government Design System](https://digitalnsw.github.io/nsw-design-system/core/colour/index.html) is available, including the subset that makes up the Department of Education brand palette. 

To list available colours by name, use `list_doestyle_colours()`:

```{r}
list_doestyle_colours()
```

More detailed information about each colour is available in the built-in dataset `colourdata`, which can be filtered for Department of Education primary brand colours, auxiliary colours, or all NSW colours.

```{r}
colourdata |>
  filter(doe_primary)
```

The available colours can also be visualised with `show_colours()`.

```{r fig-show-colours}
# Show swatches for the department's brand colours:
show_colours("brand")
```

To look up hex values, use `doe_colours()`:

```{r}
# Return hex values for named NSW Government colours:
doe_colours("blue-01", "red-02")
```

These values can be used in base R plotting functions, if you are not a ggplot2 user:

```{r fig-base-hist}
hist(public_schools$latest_year_enrolment_FTE, col = doe_colours("blue-01"))
```

### Theming

`doestyle` currently contains one ggplot2 theme, `theme_doe()`. It uses minimal non-data ink, and sets the default font to Public Sans, the department's primary brand font face. The theme is not currently very attractive - but you can help make it better by contributing to `doestyle`!

```{r fig-theme-doe}
public_schools |>
  group_by(year = year(Date_1st_teacher)) |>
  count() |>
  ggplot(aes(x = year, y = n)) +
  geom_step() +
  labs(x = "Year", y = "Government schools opened") +
  # Apply a minimalist theme that uses the Public Sans typeface
  theme_doe()
```

### Palettes

Several palettes are provided by `doestyle` in a list of lists called `doe_palettes`. If used with no `palette` argument, `scale_colour_doe()` and `scale_fill_doe()` default to the values provided in `doe_palettes$default`:

```{r}
doe_palettes$default
```

See `help(doe_palettes)` for more information on the built-in palettes.

Custom palettes are easy to create with the assistance of `pal_colour()` from the [`palettes`](https://mccarthy-m-g.github.io/palettes/) package.

```{r fig-custom-palette}
# Load `palettes`
library(palettes)

# Define a custom palette, using named NSW Government colours.
my_custom_doe_palette <- pal_colour(
  doe_colours("blue-02", "grey-02", "red-02"))

public_schools |>
  ggplot(aes(x = Date_1st_teacher,
             y = latest_year_enrolment_FTE,
             colour = School_gender,
             alpha = School_gender)) +
  geom_point() +
  scale_alpha_manual(values =  c(1, 0.1, 1)) +
  # Apply our custom palette to the `colour` aesthetic. Note that `colour` has
  # already been mapped onto the value of `School_gender` in this plot's
  # aesthetic mapping.
  scale_colour_doe(my_custom_doe_palette)
```

